---
title: "Nanobodies Sequencing Analysis"
author: 
  - name: "Francesco Lescai"
    affiliation: "NIBSC"
    email: "francesco.lescai@nibsc.org"
  - name: "Thomas Bleazard"
    affiliation: "NIBSC"
    email: "thomas.bleazard@nibsc.org"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: nibscdoc::report_html
editor_options:
  chunk_output_type: console
params:
  clusterList: NULL
  histoList: NULL
  tableList: NULL
  sampleData: NULL
---

```{R, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
library(knitr)
library(ggplot2)
library(tidyverse)
library(kableExtra)
library(ggtree)
library(Biostrings)
opts_chunk$set(message=FALSE, error=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE), cache=TRUE, cache.lazy = FALSE, echo = FALSE,  results = 'asis', fig.pos='H', fig.wide=TRUE)
```

```{R dataprep}
clusterfiles <- strsplit(params$clusterList, ",")
histofiles <- strsplit(params$histoList, ",")
tablefiles <- strsplit(params$tableList, ",")
sampleMetaData <- read_tsv(params$sampleData)
```


# Introduction



# Results 

## Clusters


We review the distribution of members of clusters and their sequence identity to the representative.



```{R clustermembership}
clusterbig <- data.frame()
for (file in clusterfiles) {
  samplename <- gsub("_clusters.summary", "", file)
  clusterdata <- read_csv(file)
  clusterdata$Sample <- rep(samplename, times=nrow(clusterdata))
  clusterbig <- rbind(clusterbig, clusterdata)
  rm(clusterdata)
}
```


```{R clusterSizePlot}
ggplot(data=clusterbig) +
    geom_histogram(aes(x=Count)) +
    scale_x_continuous(limits = c(0, 500)) +
    scale_y_continuous(limits=c(0,5000)) +
    xlab("Number of Cluster Members") +
    ggtitle("Cluster Sizes")+
  facet_wrap(~Sample)
```


```{R clusterMembersPlot}
ggplot(data=clusterbig) +
    geom_point(aes(x=Count, y=Identity), position="jitter") +
    xlab("Number of Cluster Members") +
    ylab("Average Sequence Identity\nto Cluster Representative")+
  facet_wrap(~Sample)
```



We summarise cluster counts based on the number of member sequences.

```{R clustercountsTable}
clusterbig$five <- clusterbig$Count >=5
clusterbig$hundred <- clusterbig$Count >= 100
clusterbig$thousand <- clusterbig$Count >= 1000
clustercounts <- clusterbig %>%
  group_by(Sample) %>%
  summarize(Clusters = n(), Clusters_of_5 = sum(five), Clusters_of_100 = sum(hundred), Clusters_of_1000 = sum(thousand))
kable(clustercounts) %>%
  kable_styling()
```



## CDR3 analysis 




We use a Python script to collect the CDR3 sequences only from the full amino acid sequence, using the context of the CDR3 which begins with amino acids which commence after a YYC and terminates with the amino acids which precede a WGQ. We also compute the length distribution of these CDR3 sequences.

Note that this approach is not robust to amino acid sequences which do not follow this pattern, or contain the motifs in other positions.



```{R cdrhistogram}
cdrhists <- data.frame()
for (file in histofiles) {
  samplename <- gsub("_cdr3.hist", "", file)
  histdata <- read.csv(file, header=TRUE)
  histdata$Sample <- rep(samplename, times=nrow(histdata))
  cdrhists <- rbind(cdrhists, histdata)
}
```


```{r CDR3countsTable}
cdrcounts <- cdrhists %>%
  group_by(Sample) %>%
  summarize(Unique_CDR3s = sum(Count0))
kable(cdrcounts) %>%
  kable_styling()
```



```{r CDR3sizeplot}
ggplot(data=cdrhists) +
  geom_histogram(aes(x=Size, y=Count0), stat="identity") +
  facet_wrap(~Sample) +
  ggtitle("Size Distribution of Unique CDR3s")
```


## Evolution of CDR3 




```{r fastaRep}
fastaSeq <- data.frame()
for (fasta in tablefiles){
  sample <- gsub("_cdr3.tsv", "", fasta)
  seqs <- read_tsv(fasta)
  fastaSeq <- rbind(
    fastaSeq,
    cbind(seqs,
          sample = sample)
  )
  rm(seqs)
}
```


```{r mergeData}
boostEvol <- clusterbig %>%
  left_join(fastaSeq,
            by = c(
              "Representative" = "ID",
              "Sample" = "sample"
            ))
```

temporarily taken from merged dataset, otherwise would be loaded from file.

```{r sampleMetaData}
boostEvol <- boostEvol %>%
  left_join(sampleMetaData,
      by = c("Sample" = "ID")
    )
write_tsv(boostEvol, "cdr3_boost_overview_table.tsv")
```


```{R saveImage}
save.image("nanobodies_report.RData")
```




# Additional QC


```{r viewlengthdist, eval=FALSE}
for (samplename in c("317-D1-Bst3_S1","317-D1-Bst4_S2","317-L1-Bst2_S3","317-L1-Bst3_S4","317-L1-Bst4_S5")) {
  lengthfreqs <- read.csv(file=paste0(samplename, "_L001.hist"), sep="\t", header=FALSE)
  colnames(lengthfreqs) <- c("sequence_length", "count")
  print(ggplot(data=lengthfreqs) +
    geom_histogram(aes(x=sequence_length, y=count), stat="identity") +
    ggtitle(paste0("Distribution of merged sequence lengths for sample ",samplename)))
}
```



# Methods

